Analyzing Fairness: Race, Recidivism, and the COMPAS Algorithm

### 2022-04-19 final_metrics_analysis.Rmd
Mosaic of charge degree by race

```{r py setup, include=FALSE}
library(reticulate)
#use_condaenv("anaconda3")
```

@@

import matplotlib.pyplot as plt
import matplotlib.patches as mpatches

@@

>>>>

```{r py setup, include=FALSE}
library(reticulate)

# for Dianne
use_condaenv("base")

# for Amrita
#use_condaenv("anaconda3")
```

@@

import matplotlib.pyplot as plt
import matplotlib.patches as mpatches
from statsmodels.graphics.mosaicplot import mosaic
from itertools import product

@@

```{python crosstable of charge degrees by race}
crosstable=pd.crosstab(df_viz['race'],df_viz['c_charge_degree'])
crosstable
```

```{python mosiac of charge degrees by race}
props={}
props[('Caucasian','F')]={'facecolor':'red', 'edgecolor':'white'}
props[('Caucasian','M')]={'facecolor':'red', 'edgecolor':'white'}
props[('African-American','F')]={'facecolor':'xkcd:aqua','edgecolor':'white'}
props[('African-American','M')]=        {'facecolor':'xkcd:aqua','edgecolor':'white'}
labelizer=lambda k:{('Caucasian','F'):1242,('African-American','F'):2194,('Caucasian','M'):858,('African-American','M'):979}[k]
mosaic(df_viz,['race','c_charge_degree'],labelizer=labelizer,properties=props)
plt.show()
```

59.14% of white defendants were charged with felonies versus 69.15% of black defendants. Likewise, a higher percentage of white defendants (40.86%) were charged with a misdemeanor than black defendants (30.85%).
